---
title: "Introduction to LongTermPAM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to LongTermPAM}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
# This chunk only includes setting for this R markdown, please do not run scripts in this chunk
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
) 
knitr::opts_chunk$set(eval = FALSE)
```

# Objective

This file includes the detailed instruction of **how to use LongTermPAM
R package to clean or filter long-term Monitoring PAM (MONI-PAM)
chlorophyll fluorescence (ChlF) data and calculate ChlF parameters**.
The background of this R package, such as filter logical and parameter
calculations, can be found in Zhang  et al., (2025 in submission).

# Install LongTermPAM package using:

```{r}
# Note: The # symbol is used for commenting, not for actual code execution. To run the following code, simply remove # at the beginning.
#remotes::install_github("chaoxzhang/LongTermPAM", build_vignettes = TRUE)
```

# Example datasets

**To test the code with an example data-set**, you can run this
Rmarkdown file ("Intro_To_LongTermPAM.Rmd"). First download it to a
local directory, the file is located here:

This introduction includes **two example datasets** recorded at the
Hyytiälä field station in Finland: one from the 2014–2015 observation
period and another from 2016–2017.

**Sections 1 to 3** cover data processing for the 2014–2015 dataset.
**Section 4** focuses on data processing for the 2016–2017 dataset.

**The example dataset is located here:**


Once you have installed the LongTermPAM R package on your computer,
these example datasets will be included and can be accessed using the
following codes:

```{r}
#' # check the data folder/directory location measured in 2014-2015 observation season
data20142015.path<-system.file('extdata/2014-2015',package='LongTermPAM')
#' # list all the files inside 2014-2015 folder
list.files(data20142015.path)
#'
#' # check the data folder location measured in 2016-2017 observation season
 data20162017.path<-system.file('extdata/2016-2017',package='LongTermPAM')
#' # list all the files inside 2016-2017 folder
 list.files(data20162017.path)
#'
#' #MONI-PAM data can be checked by using basic R functions, for example
 filename<-list.files(data20162017.path,full.names=T)
#' #check first MONI-PAM file inside 2016-2017 folder
 test1<-read.delim(filename[1])
# open the file by using such as View(test1)
 View(test1)
```

# Test your own data

**To test the code and clean your own dataset**, download the R Markdown
file Intro_To_LongTermPAM.Rmd to your local directory. You can then
replace the example dataset with your own data while using this R
Markdown file.

# load R packages that needed for this Rmarkdown file

```{r}
# load LongTermPAM package
library(LongTermPAM)
#load other packages need used for this file
library(data.table)
library(dplyr)
library(lubridate)
# your can use getwd function to check the current work directory location
getwd() 
```

# 1. organize and pre-process MONI-PAM data

**Before data processing and filtering, let's create the directory to
save the output data and figures.**

```{r}
# This is only an example, you can decide where the cleaned/filtered data and visualized figure will be saved 

# Function to create a directory if it doesn't exist
create_dir_if_not_exists <- function(dir_name) {
  if (!dir.exists(dir_name)) {
    dir.create(dir_name, recursive = TRUE)  # recursive = TRUE ensures that any necessary parent directories are created
  }
}

# Create a folder for saving the organized data, i.e., all the output data from Section 1.
create_dir_if_not_exists('reorganized_data')

# Create a folder for saving the output data, i.e., all the output data from Sections 2 and 3
create_dir_if_not_exists('saved_data')

# Create folders to save different data visualization figures
create_dir_if_not_exists('figure1_before_filter')
create_dir_if_not_exists('figure2_during_filter')
create_dir_if_not_exists('figure3_FmR_F0R')
create_dir_if_not_exists('figure4_diurnal_parameters')
create_dir_if_not_exists('figure5_seasonal_parameters')
```

## 1.1 Read and organize MONIPAM data: readPAM function

The function **readPAM** is used to read and organize raw MONI-PAM files

```{r}
PAM2014.2015<-
  readPAM(source.path=system.file("extdata/2014-2015/",package = "LongTermPAM"),
               pam.pattern = '.PAM|.pam',
               save.path = 'reorganized_data',# it is better to modify this to your local path such as 'c:/reorganized_data'
               site.lat =61.84562, # hyytiala latitude
               site.lon = 24.29077, # hyytiala longitude
               local.tz = 'EET', # time zone of Finland
               tz.summer =  3,
               tz.winter = 2,
               measure.time = 'winter' # In Hyytiala forest station, we always use winter time
               )
```

The readPAM() function reorganizes and combines '.PAM' raw data files
from a single observation year/folder. The function outputs a new '.csv'
file for each '.PAM' file. The function also generates three additional
combined type '.dat' files:

-   **rawPAM_2014_2015.dat** This is combined original PAM data.

-   **preprocesPAM_2014_2015.dat**\
    This file combined all the organized data and this is the file used
    for further data processing and cleaning. In this file, columns
    "plot.group", "dateBack12h", "datetimeBack12h", "sunrise",
    "sunriseEnd", "solarNoon", "sunsetStart","sunset", and "dusk" were
    added for data cleaning and data visualization purposes.

-   **preprocesPAM_head_timerange_2014_2015.dat** this file only
    contains 4 columns:

    -   "head": MONI-HEAD number
    -   "tree_num"
    -   "start.date"(start measuring time of this MONI-head and
        tree_num)
    -   "end.date" (end measuring time of this MONI-head and tree_num)

By checking this file, we can easily identify mistakes in head and/or
tree_num columns. If there is a mistake, we need to modify the head
and/or tree_num information in the **preprocesPAM_2014_2015.dat** file
to ensure correct data management, cleaning, and analysis.

## 1.2 Correct head and/or tree_num information if needed

It is important to always check the head and tree_num identifying
information for correctness (e.g., human errors when typing names in the
field). In our example we do this by reading
**preprocesPAM_head_timerange_2014_2015.dat**.

```{r}
#read the head information into R, remember where you saved your data, i.e., save.path setting when running readPAM function.
head.timerange<-read.csv("reorganized_data/preprocesPAM_head_timerange_2014_2015.dat",header = T,sep = ';')
#print this head and tree information to check the correctness
head.timerange
```

In this example, we can see that after 2015-08-28 06:03:45, all the
tree_num were given wrong names by the instrument operator. Therefore,
we need to correct them in the combined and well organized file
**preprocesPAM_2014_2015.dat** accordingly.

Note that the time format (in the start.data and end.data columns)
matches the original measurement setting and is typically consistent
across all other measurements at the same study site. In this example,
winter time zone was used for the observation period in Finland.

```{r}
# read preprocesPAM_2014_2015.dat into R using fread function from data.table package
preproces.20142015<-fread('reorganized_data/preprocesPAM_2014_2015.dat')

# Use the formatPAMdata function from this package to arrange the preproces.20142015 data into correct data format
preproces.20142015<-formatPAMdata(preproces.20142015)

# correct tree_num by using case_when function
preproces.20142015<-
  preproces.20142015 %>%
  mutate(tree_num=case_when(
    head==6~'T7TOP',
    head==11~'T7LOW',
    head==14~'T6LOW',
    head==22~'T5TOP',
    head==25~'T6TOP'
  )) %>%
  mutate(tree_num=as.factor(tree_num)) %>% 
  mutate(head_tree=paste0(head,'_',tree_num))

# check if the tree_num column is correct now
levels(preproces.20142015$tree_num)
# check first few lines of the current data
head(preproces.20142015)
```

## 1.3 Remove first few days data

We may wish to remove the first few days of data, as usually after
inserting the leaves/needles into MONI-PAM clips, the leaves/needles
need some time to adjust themselves to new growing conditions.

NOTE that You can check the original data using WinControl software
before deciding which days you want to remove.

**Here we remove the data before 2014-08-28 12:00 after checking the
data using WinControl software**

```{r}
preproces.20142015<-
  preproces.20142015 %>% 
  subset(datetime>ymd_hms('2014-08-28 12:00:00'))

# check the data now
View(preproces.20142015)
```

To avoid having to do this again, we recommend you save the modified
data into a new file but also keep this original combined file
**preprocesPAM_2014_2015.dat** without further modification.

## 1.4 Correct F for temperature dependency

The MONI-PAM instrument has a temperature dependency which can be
corrected. The correction parameters were given default value but can
also be adjusted. Details check helps of correctF() function.

**Note that this is an optional step, the user can decide if there is
need to do fluorescence correction by using this function.**

```{r}
preproces.20142015.correctF<-correctF(preproces.20142015)
```

## 1.5 Save the modified/corrected file

```{r}
# use write.table function to save the modified/corrected file to local directory
write.table(preproces.20142015.correctF,file='reorganized_data/corrected2014-2015.dat',row.names = F,sep = ';')
```

# 2. Data cleaning

## 2.1. Read organized and corrected MONI-PAM data for data cleaning

The following codes show how to filer the MONI-PAM data due to various
artefacts caused by such as snow and condensation.

```{r}
# use fread function from data.table package to read organized MONI-PAM data
flag0<-fread("reorganized_data/corrected2014-2015.dat",stringsAsFactors = T)

# First, let's see how FV/FM from original data varies across entire observing season.
# We can calculate FV/FM using FindFVFM function, visualize FV/FM data using plotFVFM function
flag0.fvfm<-  FindFVFM(PAM.data=flag0,save.path = 'saved_data',
                       save.title = 'Flag 0',save.file=T)

# plot and save FVFM figure to local folder
plotFVFM(flag0.fvfm, plot.title = 'Flag 0 fvfm',save.path='figure5_seasonal_parameters')

# visualize data before filtering using plotPAM function 
plotPAM(flag0, plot.title = 'Before filter',save.path = 'figure1_before_filter')
```

**Note: Filtering must follow the sequential order from step 1 to step
6. Each step depends on the output data of the previous one (e.g.,
filter 2 uses filtered data from filter 1, filter 3 from filter 2,
etc.). Skipping steps is not allowed.**

**cleaning data step 1**

```{r}
# filter1.lowF
flag1.lowF<-filter1.lowF(PAM.data=flag0,save.file=T,save.path= 'saved_data')
#check FVFM after filter step 1
flag1.fvfm<-FindFVFM(flag1.lowF,save.file=T, save.path ='saved_data',
                     save.title = 'Flag 1')
#plot and save FVFM figure to local folder
plotFVFM(flag1.fvfm,plot.title = 'Flag 1 fvfm',
         save.path='figure5_seasonal_parameters')
#visualize data filtering
plotCheckFilter(dataAfterFilter=flag1.lowF,dataBeforeFilter=flag0,
                save.path='figure2_during_filter')
```

**continue for steps 2 to 6**

For steps 2 to 6, you must set various threshold parameters that affect
filter performance. Detailed about the meanings of these parameters or
arguments can be found in R help page and in the paper Zhang ,
(2015, in submission). You need to carefully (visually) check the
cleaned data (figure output from **plotCheckFilter** function). This is
why there is always **plotFVFM** and **plotCheckFilter** function in the
end of each filter step to visualize your data filtering results.

Note that **plotFVFM** checks whole growing season data filtering
results and **plotCheckFilter** checks day by day.

**Refine these parameters until you feel your filter is good enough by
checking plotted diurnal figures, then move to the next filter step:**

```{r}
# adjust f2.Fm from 0.01 to 0.1 by a 0.01 interval
# fmYII = 3 usually is a good setting, you can just try to adjust f2.Fm
flag2.night<-
   filter2.night(flag1.lowF,f2.Fm=0.05,f2.FmYII=3,
                 save.path='saved_data',save.file = T)
flag2.fvfm<-
  FindFVFM(flag2.night,save.file=T, save.path ='saved_data',save.title = 'Flag 2')
plotFVFM(flag2.fvfm,plot.title = 'Flag 2 fvfm',
         save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter=flag2.night,dataBeforeFilter = flag0,
                save.path='figure2_during_filter')
```

```{r}
# adjust f3.YII from 0.01 to 0.1 by a 0.01 interval
# adjust f3.Fm from 0.05 to 0.3 by a 0.05 interval
# fmYII = 3 usually is a good setting, you can just try to adjust f3.YII and f3.Fm 
flag3.day<-
   filter3.day(flag2.night,f3.YII=0.02,f3.Fm=0.1,f3.FmYII = 3,
               save.file = T,save.path ='saved_data')
flag3.fvfm<-
  FindFVFM(flag3.day,save.file=T, save.path ='saved_data',save.title = 'Flag 3')

plotFVFM(flag3.fvfm,plot.title = 'Flag 3 fvfm',save.path='figure5_seasonal_parameters')

plotCheckFilter(dataAfterFilter = flag3.day,dataBeforeFilter =flag0,
                save.path='figure2_during_filter')
```

```{r}
# adjust f4.Fm from 0.05 to 0.3 by a 0.05 interval
# f4.FmYII = 3 usually is a good setting, you can just try to adjust f4.Fm
flag4.FvFm<-
   filter4.FVFM(flag3.day,save.file = T,f4.Fm=0.15,f4.FmYII=3,
                    save.path = 'saved_data')
flag4.fvfm<-
  FindFVFM(flag4.FvFm,save.file=T,save.path = 'saved_data',save.title = 'Flag 4')

plotFVFM(flag4.fvfm,plot.title = 'Flag 4 fvfm',save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter = flag4.FvFm,dataBeforeFilter = flag0,
                save.path='figure2_during_filter')
```

```{r}
# adjust f5.Fm from 0.05 to 0.3 by a 0.05 interval
# f5.FmYII = 3 usually is a good setting, you can just try to adjust f5.Fm
flag5.expand<-filter5.expand(flag4.FvFm,f5.Fm = 0.2,f5.FmYII = 3,
                             save.file = T, save.path = 'saved_data')
flag5.fvfm<-
  FindFVFM(flag5.expand,save.file=T,save.path='saved_data',save.title = 'Flag 5')

plotFVFM(flag5.fvfm,plot.title = 'Flag 5 fvfm',save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter = flag5.expand,dataBeforeFilter = flag0,
                save.path='figure2_during_filter')
```

```{r}
# adjust expand.time  according your data recording interval, 60 mins here for this example include 2 measurement points
flag6.adjacent<-filter6.adjacent(flag5.expand,expand.time = 60,
                                 save.file = T, save.path = 'saved_data')
flag6.fvfm<-
  FindFVFM(flag6.adjacent,save.file=T,save.path = 'saved_data',
           save.title = 'Flag 6')

plotFVFM(flag6.fvfm,plot.title = 'Flag 6 fvfm',
         save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter=flag6.adjacent,dataBeforeFilter=flag0,
                save.path='figure2_during_filter')
```

# 3. Calculate other paramters

## 3.1 Estimate FmR and FoR using ChlFRef function

To calculate additional chlorophyll fluorescence parameters, Fm
reference (FmR) and Fo reference (FoR) are needed. See **ChlFRef**
function help and Zhang et al., 2015 in submission for detailed
estimation instructions.

The output of ChlFRef function include:

(1) A data.frame with FmR and FoR for each measuring head, along with
    model summary statistics (R2, bias,RMSE,and RRMSE) from non-linear
    regression model between Fm and Fv/Fm over the entire observation
    period.

(2) Correlation figures between Fm and Fv/Fm.

**NOTE**: You can also define FmR and FoR manually instead of using
ChlFRef. If using this package for further other ChlF parameter
calculation, your FmR dataset should be a data.frame or data.table
format, including the columns:
'season','head_tree','head','tree_num','FmR' and 'FoR'.

```{r fig.height=6,fig.width=20}
FmR.df<-ChlFRef(fvfm.beforeFilter = flag0.fvfm,
                fvfm.afterFilter = flag6.fvfm,
                FvFmR = 0.83,save.file = T, 
                save.path = 'figure3_FmR_F0R')
FmR.df
```

## 3.2 Estimate diurnal parameters using diurnalParams function

This function estimates diurnal parameters,including:

(1) quenching parameters: PQ, qL~T~, NPQ~T~, and NPQr;

(2) quantum yield parameters: $\Phi$P [or Y(II)], $\Phi$NPQ~T~,
    $\Phi$f,D, and $\Phi$NPQr .
    
```{r}
diurn.para<-diurnalParams(filtered.data = flag6.adjacent,
                          filtered.fvfm = flag6.fvfm,
                          FmR.data = FmR.df,
                          save.file = T,save.path = 'saved_data')
```

## 3.3 Estimate seasonal parameters using seasonalParams function

This function estimates seasonal parameters,including:

(1) quenching parameters: PQ~S~, qL~S~, and NPQ~S~;

(2) quantum yield parameters: F~V~/F~M~, $\Phi$NPQ~S~, and $\Phi$f,D~S~.
  
```{r}
seas.para<- seasonalParams(fvfm.afterFilter = flag6.fvfm,
                           FmR.data = FmR.df,
                           save.file = T,save.path = 'saved_data')  
```

## 3.3 Visualize diurnal and seasonal parameters

plot seasonal parameters and save using the below function:

```{r fig.height=3.5,fig.width=15}
plotSeasonPara(season.param = seas.para,save.path = 'figure5_seasonal_parameters') 
```

plot diurnal parameters and save using the below function:

```{r fig.height=10,fig.width=15}
# plot and save diurnal quenching parameters: PQ,qLT,NPQT, and NPQr, along with PAR and temperature
plotDiurnQuench(diurnal.para = diurn.para,save.path = 'figure4_diurnal_parameters')

# plot and save quantum yield parameters: Phi_P(YII),Phi_NPQT,Phi_fD, and Phi_NPQr
plotDiurnYield(diurnal.para = diurn.para,save.path = 'figure4_diurnal_parameters')
```

# 4. Additional example from 2016-2017

Here we repeat steps 1-3 using data from a different year.

**Before data processing and filtering, let's create the directory to
save the output data and figures.**

```{r}
# This is only an example, you can decide where the cleaned/filtered data and visualized figure will be saved 

# Function to create a directory if it doesn't exist
create_dir_if_not_exists <- function(dir_name) {
  if (!dir.exists(dir_name)) {
    dir.create(dir_name, recursive = TRUE)  # recursive = TRUE ensures that any necessary parent directories are created
  }
}

# Create a folder for saving the organized data, i.e., all the output data from Section 1.
create_dir_if_not_exists('reorganized_data')

# Create a folder for saving the output data, i.e., all the output data from Sections 2 and 3
create_dir_if_not_exists('saved_data')

# Create folders to save different data visualization figures
create_dir_if_not_exists('figure1_before_filter')
create_dir_if_not_exists('figure2_during_filter')
create_dir_if_not_exists('figure3_FmR_F0R')
create_dir_if_not_exists('figure4_diurnal_parameters')
create_dir_if_not_exists('figure5_seasonal_parameters')
```

## 4.1 Read and organize MONI-PAM data: readPAM function

```{r}
PAM2016.2017<-
  readPAM(source.path=system.file("extdata/2016-2017/",package = "LongTermPAM"),
               pam.pattern = '.PAM|.pam',
               save.path = 'reorganized_data',# it is better to modify this to your local path
               site.lat =61.84562, # hyytiala latitude
               site.lon = 24.29077, # hyytiala longitude
               local.tz = 'EET', # time zone of Finland
               tz.summer =  3,
               tz.winter = 2,
               measure.time = 'winter' # In Hyytiala forest station, we always use winter time
               )
```

## 4.2 Correct head and/or tree_num information if needed

```{r}
#read the head information into R, remember your saved-data path
head.timerange<-read.csv("reorganized_data/preprocesPAM_head_timerange_2016_2017.dat",header = T,sep = ';')
#print this head and tree information
head.timerange
```

In this example, we identify (1) two unknown tree_num entries: Unknow1
and Unknow2, measured for only two days, and (2) head 25 with tree_num
'Pine2LOW' which contains very few data lines. These three trees are
unnecessary and can be removed form the preprocessed dataset
**preprocesPAM_2016_2017.dat** (but not from this head info file)

```{r}
# read preprocesPAM_2016_2017.dat into R using fread from data.table package
preproces.20162017<-fread('reorganized_data/preprocesPAM_2016_2017.dat')

# arrange the preproces.20162017 data to correct data format using formatPAMdata function from this package
preproces.20162017<-formatPAMdata(preproces.20162017)

# correct tree_num
preproces.20162017<-
  preproces.20162017 %>% 
  subset(tree_num!='Unknow1'&tree_num!='Unknow2'&tree_num!='Pine2LOW') %>% 
  droplevels() %>% 
  mutate(head_tree=paste0(head,'_',tree_num))

# check tree_num
levels(preproces.20162017$tree_num) # we can see now 'Unknow1', 'Unknow2' and 'Pine2LOW' are removed
# check data
head(preproces.20162017)
```

## 4.3 Remove first few days data

**Here we remove the data before 2016-11-05 12:00 after checking the
data using WinControl software**

```{r}
preproces.20162017<-preproces.20162017 %>% subset(datetime>ymd_hms('2016-11-05 12:00:00'))
# check the data now
head(preproces.20162017)
```

## 4.4 Correct F for temperature dependency

The correction parameters were given default values, but can be adjusted
for the specific instrument. Details check helps of correctF() function

```{r}
preproces.20162017.correctF<-correctF(preproces.20162017)
```

## 4.5 Save the modified/corrected file

```{r}
# write.table function from R to save data into local folder
write.table(preproces.20162017.correctF,file='reorganized_data/corrected2016-2017.dat',
            row.names = F,sep = ';')
```

## 4.6 Data cleaning

```{r}
# use fread function from data.table package to read organized MONI-PAM data
flag0<-fread("reorganized_data/corrected2016-2017.dat",stringsAsFactors = T)

# First, let's see how FV/FM from original data varies across entire observing season.
# We can calculate FV/FM using FindFVFM function, visualize FV/FM data using plotFVFM function
flag0.fvfm<-  FindFVFM(PAM.data=flag0,save.path = 'saved_data',
                       save.title = 'Flag 0',save.file=T)

# plot and save FVFM figure to local folder
plotFVFM(flag0.fvfm, plot.title = 'Flag 0 fvfm',save.path='figure5_seasonal_parameters')

# visualize data before filtering using plotPAM function 
plotPAM(flag0, plot.title = 'Before filter',save.path = 'figure1_before_filter')
```

```{r}
################ filter step 1 ############################
# filter1.lowF
flag1.lowF<-filter1.lowF(PAM.data=flag0,save.file=T,save.path= 'saved_data')
#check FVFM after filter step 1
flag1.fvfm<-FindFVFM(flag1.lowF,save.file=T, save.path ='saved_data',
                     save.title = 'Flag 1')
#plot and save FVFM figure to local folder
plotFVFM(flag1.fvfm,plot.title = 'Flag 1 fvfm',
         save.path='figure5_seasonal_parameters')
#visualize data filtering
plotCheckFilter(dataAfterFilter=flag1.lowF,dataBeforeFilter=flag0,
                save.path='figure2_during_filter')

################ filter step 2 ############################
# adjust f2.Fm from 0.01 to 0.1 by a 0.01 interval
# fmYII = 3 usually is a good setting, you can just try to adjust f2.Fm
flag2.night<-
   filter2.night(flag1.lowF,f2.Fm=0.05,f2.FmYII=3,
                 save.path='saved_data',save.file = T)
flag2.fvfm<-
  FindFVFM(flag2.night,save.file=T, save.path ='saved_data',save.title = 'Flag 2')
plotFVFM(flag2.fvfm,plot.title = 'Flag 2 fvfm',save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter=flag2.night,dataBeforeFilter = flag0,
                save.path='figure2_during_filter')

################ filter step 3 ############################
# adjust f3.YII from 0.01 to 0.1 by a 0.01 interval
# adjust f3.Fm from 0.05 to 0.3 by a 0.05 interval
# fmYII = 3 usually is a good setting, you can just try to adjust f3.YII and f3.Fm 
flag3.day<-
   filter3.day(flag2.night,f3.YII=0.02,f3.Fm=0.1,f3.FmYII = 3,
               save.file = T,save.path ='saved_data')
flag3.fvfm<-
  FindFVFM(flag3.day,save.file=T, save.path ='saved_data',save.title = 'Flag 3')

plotFVFM(flag3.fvfm,plot.title = 'Flag 3 fvfm',save.path='figure5_seasonal_parameters')

plotCheckFilter(dataAfterFilter = flag3.day,dataBeforeFilter =flag0,
                save.path='figure2_during_filter')

################ filter step 4 ############################
# adjust f4.Fm from 0.05 to 0.3 by a 0.05 interval
# f4.FmYII = 3 usually is a good setting, you can just try to adjust f4.Fm
flag4.FvFm<-
   filter4.FVFM(flag3.day,save.file = T,f4.Fm=0.15,f4.FmYII=3,
                    save.path = 'saved_data')
flag4.fvfm<-
  FindFVFM(flag4.FvFm,save.file=T,save.path = 'saved_data',save.title = 'Flag 4')

plotFVFM(flag4.fvfm,plot.title = 'Flag 4 fvfm',save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter = flag4.FvFm,dataBeforeFilter = flag0,
                save.path='figure2_during_filter')

################ filter step 5 ############################
# adjust f5.Fm from 0.05 to 0.3 by a 0.05 interval
# f5.FmYII = 3 usually is a good setting, you can just try to adjust f5.Fm
flag5.expand<-filter5.expand(flag4.FvFm,f5.Fm = 0.2,f5.FmYII = 3,
                             save.file = T, save.path = 'saved_data')
flag5.fvfm<-
  FindFVFM(flag5.expand,save.file=T,save.path='saved_data',save.title = 'Flag 5')


plotFVFM(flag5.fvfm,plot.title = 'Flag 5 fvfm',save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter = flag5.expand,dataBeforeFilter = flag0,
                save.path='figure2_during_filter')

################ filter step 6 ############################
# adjust expand.time  according your data recording interval, 60 mins here for this example include 2 measurement points
flag6.adjacent<-filter6.adjacent(flag5.expand,expand.time = 60,
                                 save.file = T, save.path = 'saved_data')
flag6.fvfm<-
  FindFVFM(flag6.adjacent,save.file=T,save.path = 'saved_data',
           save.title = 'Flag 6')

plotFVFM(flag6.fvfm,plot.title = 'Flag 6 fvfm',
         save.path='figure5_seasonal_parameters')
plotCheckFilter(dataAfterFilter=flag6.adjacent,dataBeforeFilter=flag0,
                save.path='figure2_during_filter')
```

## 4.7 Calculate other paramters

**Estimate FmR and FoR using ChlFRef function**

```{r fig.height=6,fig.width=20}
FmR.df<-ChlFRef(fvfm.beforeFilter = flag0.fvfm,fvfm.afterFilter = flag6.fvfm,
                FvFmR = 0.83,save.file = T, save.path = 'figure3_FmR_F0R')
FmR.df
```

**Calculate additional ChlF parameters**

```{r}
# estimate diurnal parameters using diurnalParams function
diurn.para<-diurnalParams(filtered.data = flag6.adjacent,
                          filtered.fvfm = flag6.fvfm,
                          FmR.data = FmR.df,
                          save.file = T,save.path = 'saved_data')

# estimate seasonal parameters using seasonalParams function
seas.para<- seasonalParams(fvfm.afterFilter = flag6.fvfm,
                           FmR.data = FmR.df,
                           save.file = T,save.path = 'saved_data')  
```

**Visualize additional ChlF parameters**

plot seasonal parameters and save using the below function:

```{r fig.height=3.5,fig.width=15}
plotSeasonPara(season.param = seas.para,save.path = 'figure5_seasonal_parameters') 
```

plot diurnal parameters and save using the below function:

```{r fig.height=10,fig.width=15}
# plot and save diurnal quenching parameters: PQ,qLT,NPQT, and NPQr, along with PAR and temperature
plotDiurnQuench(diurnal.para = diurn.para,save.path = 'figure4_diurnal_parameters')

# plot and save quantum yield parameters: Phi_P(YII),Phi_NPQT,Phi_fD, and Phi_NPQr
plotDiurnYield(diurnal.para = diurn.para,save.path = 'figure4_diurnal_parameters')
```

